# Data 607 - Regex and string functions assignment
### Heather Geiger
### February 17,2018

## Part 1 - Formatting names using regular expressions

First, we get the names from the raw data.

```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- unlist(str_extract_all(raw.data,"[[:alpha:]., ]{2,}"))
names
```

Next, we convert to first_name last_name format.

To do this, we split by blank space.

Then we check for commas in the first element after split.

Finally, put whatever is before the comma last and then remove the comma.

```{r}
names_split_by_spaces <- str_split(names,"[[:blank:]]+")
names_split_by_spaces
for(i in 1:length(names))
{
first_element_before_space <- names_split_by_spaces[[i]][1]
if(str_detect(first_element_before_space,",") == TRUE){
	first_element_minus_comma <- str_replace(names_split_by_spaces[[i]][1],pattern=",",replace="")
	names_split_by_spaces[[i]] <- c(names_split_by_spaces[[i]][2:length(names_split_by_spaces[[i]])],first_element_minus_comma)
	}
}
names_split_by_spaces
```

Next, we check if there is a title, like Rev. or Dr.
For this, we look for two or more alphabetical characters followed by a ".".

```{r}
has_title <- c()

for(i in 1:length(names))
{
first_element <- names_split_by_spaces[[i]][1]
has_title <- c(has_title,str_detect(first_element,"[[:alpha:]]{2,}\\."))
}

has_title
```

To check for a second name, we check the number of elements per item in names_split_by_spaces.

If there are more than two elements (minus the title), then the person has a second name.

```{r}
second_name <- c()

for(i in 1:length(names))
{
num_elements <- length(names_split_by_spaces[[i]])
if(has_title[i] == TRUE){num_elements <- num_elements - 1}
second_name <- c(second_name,num_elements > 2)
}

second_name
```

Now that we are done checking for titles and second names, let's convert names_split_by_spaces from a list back into a vector. 

Let's put the items in each list pasted together separated by a space.

```{r}
firstname_lastname <- c()

for(i in 1:length(names))
{
firstname_lastname <- c(firstname_lastname,paste0(names_split_by_spaces[[i]],collapse=" "))
}

data.frame(Name = firstname_lastname,Has.title = has_title,Second.name = second_name)
```

## Part 2 - Finding strings that match a given regex

1. [0-9]+\\$

The double backslash here means that we are looking for a literal $. [0-9]+ means one or more digits. Together, these mean we are looking for one or more digits followed by a dollar sign.

I tested this using some example strings. The fact that the string ending in digits, but with no literal $, did not match confirms that we are looking for literal $ with this regex.

```{r}
mystrings <- c("9-to-5","seventy","85$","$$","abc7$","99")
unlist(str_extract_all(mystrings,"[0-9]+\\$"))
```

Here, 85$ and the 7$ part of abc7$ both match the regex.

2. \\b[a-z]{1,4}\\b

Here, we are looking for a word edge, followed by 1-4 alphabetical characters, followed by a word edge.

A word edge can be defined as either a word being at the beginning of a string, having only blank space before it, or having other things before it but then having a blank space between those other things and the word start.

```{r}
mystrings <- c("abc"," abc","abcde fghij","abc def gh ijkl mnopq rst uv9 wx") 
str_extract_all(mystrings,"\\b[a-z]{1,4}\\b")
```
In the first two strings, the words are both "abc". So we have a word edge on either side, and three alphabetical characters, thus fulfilling the regex.

In the third string, both words are too long (5 characters), so we do not match.

In the last string, there are 8 words. 6 of these words match the regex. mnopq does not match because it has too many alphabetical characters before the end of the word. uv9 does not match because there is a non-alphabetical character between the two word edges.

3. .*?\\.txt$

The "\\.txt$" part of this means we want the string to end in ".txt", where we want an actual "." and not just any character.

The ".*?" part means zero or more matches of any character, matched at most once.

This is likely a good regular expression to find files with suffix ".txt".

```{r}
mystrings <- c("/data/analysis/hmgeiger/Project_RNA/samples.txt","file.txt.txt","file.txt.gz")
str_extract_all(mystrings,".*?\\.txt$")
```

4. \\d{2}/\\d{2}/\\d{4}

\\d{2} means 2 digits. \\d{4} means 4 digits. The "/" is a literal slash here. So 2 digits, "/", 2 digits, "/", then 4 digits should match this regex.

```{r}
mystrings <- c("123/45/6789")
str_extract_all(mystrings,"\\d{2}/\\d{2}/\\d{4}")
```

5. 

```{r}
print("<(.+?)>.+?</\\1>")
```

No backslashes, so the first "<" and ">" are both literal characters.

So, first match "<", then one or more characters, then ">".

Then match one or more characters again, then "<", then "/".

Then, the "\\1" means we want to look for again whatever we found that matches the part in parentheses (one or more characters).

Finally, end with a ">".

```{r}
mystring <- "<tag>stuff in between</tag>"
str_extract_all(mystring,"<(.+?)>.+?</\\1>")
```

Weirdly, I thought that this should work just as well with nothing in between the two sets of <>, because of the "?".

But it seems it does not.

```{r}
mystring <- "<tag></tag>"
str_extract_all(mystring,"<(.+?)>.+?</\\1>")
```

We don't get a match if the end part (after "/") does not match the first part exactly before ending with the >.

```{r}
mystring <- "<tag>stuff in between</tag2>"
str_extract_all(mystring,"<(.+?)>.+?</\\1>")
```

We can however put whatever we want in between the first <>, as long as we put something.

Again, I thought the "?" meant we could also put nothing, but this appears not to be the case.

```{r}
mystrings <- c("<>stuff in between</>","<12345>stuffinbetween</12345>")
str_extract_all(mystrings,"<(.+?)>.+?</\\1>")
```
